Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-22 article/article.Rmd
Commit 2024-06-22 20:14

First, we evaluated the performance of the modelled $\text{GHI}_\text{ref}$ in
relation to the measured GHI for different levels of atmospheric clearness, by using
in the RTM the monthly climatological AOD and WC, less their respective standard
deviations.
<!--
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
-->
These values represent typical atmospheres in Thessaloniki with lower than average
load of aerosols and humidity, which are the main factors that attenuate the GHI,
excluding clouds.  With this approach the simulated $\text{GHI}_\text{ref}$ should be
generally greater than the measured GHI when aerosols are more abundant.  To
compensate for this, we defined the following threshold $\text{E}$ to compare the
measured $\text{GHI}$: 
<!--
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}.
-->
\begin{equation}
\text{CE} : \text{E} > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{ref} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
<!--
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.
-->
This is the criterion of our CE identification.

@@

<!-- reasoning for cretria -->
<!-- We have to note, that the differentiation of the threshold factor was needed, because
<!-- of the high irradiance values we observed, early in the morning and late in the
<!-- afternoon. 
<!-- We have confirmed, by inspecting images from the sky cam, that we have
<!-- duration of elevated irradiance, either due to the clearness of the atmosphere, or some
<!-- interferences by reflections on nearby bright surfaces. Although, these cases may
<!-- mask some of the CE events. The actual CE events, produce higher irradiance and thus
<!-- are identified as such. As a side effect, the reported OIR will be slight
<!-- underestimated for those SZAs, but the overall contribution of those cases to the
<!-- total daily energy, is minimal due the low occurrences and the lower irradiances.
-->

@@

<!-- extreme cases criterion -->
A sub-category  of the CE events that is often discussed in the relevant literature
[@Cordero2023; @Martins2022; @Yordanov2015], are the extreme cloud enhancement (ECE)
events. These are cases of CE where the measured intensity of the irradiance exceeds
the TSI at the top of the atmosphere. In this case the threshold EEEEEEEEEE:
<!--and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.-->
\begin{equation}
\text{ECE}: \text{GHI}_\text{i} > \cos(\theta) \times E_{i\odot} / r_{i}^2
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant adjusted for the actual Sun -- Earth distance, and $i$ each of the one-minute observation.

@@

```{r example-day, echo=FALSE, fig.cap="Example of cloud identification for 2019-07-11. The green line with blue symbols depicts the measured GHI in one minute steps. The red line shows the modelled threshold for the detection of CE events, which are denoted with red circles. The black curve represents the TSI at the top of the atmosphere, adjusted for the actual Sun-Earth distance and multiplied by the cosine of the cosine of SZA.", fig.pos='H'}
knitr::include_graphics("../images/example-days-18.png")
```

>>>>

First, we evaluated the performance of the modelled $\text{GHI}_\text{ref}$ in
relation to the measured GHI for different levels of atmospheric clearness, by using
in the RTM the monthly climatological AOD and WC, less their respective standard
deviations.
<!--
$\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma$ and $w_{h\text{cs}} = w_h - 1\sigma$
($\text{GHI}_\text{CSm}$).
-->
These values represent typical atmospheres in Thessaloniki with lower than average
load of aerosols and humidity, which are the main factors that attenuate the GHI,
excluding clouds.  With this approach the simulated $\text{GHI}_\text{ref}$ should be
generally greater than the measured GHI when aerosols are more abundant.  To
compensate for this, we defined the following threshold $E$ to compare the
measured $\text{GHI}$: 
<!--
($`r 100*(C4_cs_ref_ratio-1)`\,\%$)
with an additional constant offset of
$`r C4_GLB_diff_THRES`\,\text{W}/\text{m}^2$,
as described in Equation\nobreakspace\ref{eq:CE4}.
-->
\begin{equation}
\text{CE} : E > `r C4_GLB_diff_THRES` + `r C4_cs_ref_ratio` \cdot \text{GHI}_\text{ref} \,\,[\text{W}/\text{m}^2] \label{eq:CE4}
\end{equation}
<!--
where: $\text{E}$ the measured irradiance, $\text{E}_\text{CSm}$ the selected
modelled clear sky irradiance, and $i$ each of the one-minute observation.  This is
the criterion of our CE identification.
-->
This is the criterion of our CE identification.

@@

<!-- reasoning for criteria -->
<!-- We have to note, that the differentiation of the threshold factor was needed, because
<!-- of the high irradiance values we observed, early in the morning and late in the
<!-- afternoon. 
<!-- We have confirmed, by inspecting images from the sky cam, that we have
<!-- duration of elevated irradiance, either due to the clearness of the atmosphere, or some
<!-- interferences by reflections on nearby bright surfaces. Although, these cases may
<!-- mask some of the CE events. The actual CE events, produce higher irradiance and thus
<!-- are identified as such. As a side effect, the reported OIR will be slight
<!-- underestimated for those SZAs, but the overall contribution of those cases to the
<!-- total daily energy, is minimal due the low occurrences and the lower irradiances.
-->

@@

<!-- extreme cases criterion -->
A sub-category  of the CE events that is often discussed in the relevant literature
[@Cordero2023; @Martins2022; @Yordanov2015], are the extreme cloud enhancement (ECE)
events. These are cases of CE where the measured intensity of the irradiance exceeds
the TSI at the top of the atmosphere. In this case the threshold $E$ is given by:
<!--and detected according to the Equation\nobreakspace{}\ref{eq:ECE}.-->
\begin{equation}
\text{ECE}: E > \cos(\theta) \cdot E_{\odot} \frac{r^2_\text{m}} {r^2}
\label{eq:ECE}
\end{equation}
where: $\theta$ the solar zenith angle, $E_{\odot}$ the solar constant adjusted for
the actual Sun -- Earth distance $r$ and $r_\text{m}$ is the mean Sun -- Earth
distance of $1.496\times10^8\,\text{km}$.

@@

```{r example-day, echo=FALSE, fig.cap="Example of CE identification for 2019-07-11. The green line with blue symbols depicts the measured GHI in one minute steps. The red line shows the modelled threshold $E$ for the detection of CE events, which are denoted with red circles. The black curve represents the TSI at the top of the atmosphere, adjusted for the actual Sun-Earth distance and multiplied by the cosine of the cosine of SZA.", fig.pos='H'}
knitr::include_graphics("../images/example-days-18.png")
```
